---
title: "gganimate workshop"
author: "Danny Ettelson"
date: "1/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Make your static graphs work well first before you animate before trying to animate
2. animating can take a long time to render so start with simple animations first


rules for successful visuals IN THIS ORDER
1. correct
2. clear
3. looks awesome

```{r lib}

library(tidyverse)
library(ggridges)
library(gganimate)
library(ggrepel)
library(transformr)
library(gifski)


```

Get Channel Island fox population data and wranlge it

```{r data}
#gifkey?
ci_fox_pop <- read_csv("ci_fox_pop.csv")
# tidy format has year, island name, population, year

#gather population data using gather() in tidyr

fox <- ci_fox_pop %>% 
  gather(island, pop, san_miguel:san_nicolas)




```

First make a static version of graph and use facet_wrap to view by year

```{r graphs}

#to animaate, we'll use transition_state
#splits by a variable into states similarly to facet_wrap
#then it 'tweens (interpolates) between the states

ggplot(fox, aes(x = year, y = pop)) +
  geom_point(size = 3, aes(color = island)) +
  facet_wrap(~year) #split into graphs by year


```

```{r animate}

ggplot(fox, aes(x = year, y = pop)) +
  geom_point(size = 3, aes(color = island)) +
  transition_states(states = year,
                    transition_length = 1, 
                    state_length = 1,
                    wrap = FALSE) +
  ease_aes('cubic-in-out') +
  shadow_mark()




#wrap interpolates from first to last point
```

```{r animate 2}

#transition_reveal for island fox populations
#allows lines to gradually build up
#it knows to keep old data for path and polygon layers, to help follow along labels
fox_line <- ggplot(fox, aes(x = year, y = pop)) +
  geom_line(size = 1, aes(color = island)) +
  theme_dark() +
  transition_reveal(id = island, along = year) +
  ease_aes('cubic-in-out')

#rendering code
#animate(fox_line, nframes( = 100), renderer = gifski_renderer(fox_line.gif))

animate(fox_line, nframes = 100, renderer = gifski_renderer("fox_line.gif"))

```

Use starwars dataset to create animated graphics using transition_manual() and transition_layers()

```{r starwars}

#Wrangling

sw <- starwars %>% 
  filter(species == "Human" |
           species == "Droid" |
           species == "Wookiee" |
           species == "Ewok") %>% 
  mutate(species = factor(species))


sw$species <- fct_relevel(sw$species, "Ewok", "Droid", "Human", "Wookiee")
  

#transition manual makes states based on what you want if the auto doesn't make sense
# good for discrete frames where you don't want interpolation
ggplot(sw, aes(x = height, y = mass, label = name)) +
  geom_point(aes(color = species)) +
  geom_text_repel(size = 2, segment.size = 0.2) +
  labs(title = "Species: {current_frame}") +
  transition_manual(frames = species)

```

transition_layers: gradually add layers to the plot in the order defined in the code

```{r transition_layers}
#transitioned line and error
sw_size <- ggplot(sw, aes(x = height, y = mass)) +
  geom_point(aes(color = species)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth(method = "lm", se = TRUE, color = "NA") +
  transition_layers(layer_length = 1, transition_length = 2) +
  enter_fade() +
  theme_minimal()

animate(sw_size, nframes = 100, renderer = gifski_renderer("sw_size.gif"))

```

Final example: animated ridge plotusing ransition_states with abalone sizes

```{r abalone}

abalone <- read_csv("abalone.csv") %>% 
  filter(sex == "M" | sex == "F" | sex == "I") %>% 
  filter(age_years >4 & age_years <25) %>% 
  mutate(sex = fct_relevel(factor(sex), "I", "F", "M"))
  
ggplot(abalone, aes(x = length_mm, y = age_years, fill = sex)) +
  geom_density_ridges(alpha = 0.5, color = "white") +
  scale_fill_manual(values = c("purple","blue","cyan")) +
  transition_states(age_years, transition_length = 1, state_length = 1) +
  shadow_mark()

```






